An Empirical Analysis of the Role of Energy in Economic Growth

### 2014-12-29 Articles/Article_EnerEcon_2014/EconGrowth3.Rnw
Now all sigma tables are included.
Edits to sigma table interpretations.

The interpretation of elasticities of substitution 
among factors of production ($k$, $l$, and $e$) is challenging,
because a version of each elasticity of substitution appears three or four times 
among the various nests. 
For example, an elasticity of substitution between capital~($k$) and labor~($l$) 
appears in the following nests: 
($\underline{kl}$), 
($\underline{kl}$)$e$, 
($\underline{l}e$)$\underline{k}$, and 
($e\underline{k}$)$\underline{l}$. 
The following tables assist interpretation: 
C (c) indicates strong (weak) complementarity, 
S (s) indicates strong (weak) substitutability, and
$\sim\sim\sim$ ($\sim$) indicates large (small) uncertainty.

@@

% \begin{table} \caption{Elasticities of substitution (US).} \label{tab:sigma_US} 
%   \begin{center}
%     \begin{tabular}{r l c l c l c} 
%       \toprule
%          & \multicolumn{2}{c}{$k$ \& $l$} & \multicolumn{2}{c}{$k$ \& $e$} & \multicolumn{2}{c}{$l$ \& $e$} \\
%       \midrule
%       \multirow{2}{*}{$\sigma_1$} & ($\underline{kl}$) & C 
%                                         &  & 
%                                               & &  \\
%                                   & ($\underline{kl}$)$e$ & C 
%                                         & ($\underline{ek}$)$l$ & $\sim\sim\sim$ 
%                                               & ($\underline{le}$)$k$ & $\sim\sim\sim$ \\
%       \midrule
%       \multirow{3}{*}{$\sigma$} & &
%                                       & ($\underline{k}l$)$\underline{e}$ & $\sim\sim\sim$
%                                             & ($k\underline{l}$)$\underline{e}$ & $\sim\sim\sim$ \\
%                                 & ($\underline{l}e$)$\underline{k}$ & C
%                                       & ($l\underline{e}$)$\underline{k}$ & C
%                                             &  & \\
%                                 & ($e\underline{k}$)$\underline{l}$ & C
%                                       & &
%                                             & ($\underline{e}k$)$\underline{l}$ & C \\
%       \midrule
%                                 & \multicolumn{2}{c}{C} & \multicolumn{2}{c}{C} & \multicolumn{2}{c}{C} \\
%       \bottomrule
%     \end{tabular}
%   \end{center}
% \end{table}

@@

% label is tab:sigma_SA
\eostable{SA}
          {\firstrow{c$\sim$}{$\sim\sim\sim$}{C}{$\sim\sim\sim$}}
          {\secondrow{C}{C}{C}{C}{$\sim\sim\sim$}{C\tablefootnote{Points on right are all same.}}}
          {\thirdrow{c}{C}{C}}

**** For the time being, I've added two of these tables, US and UK.
We can add the remaining seven tables easily, either here or in an appendix. 
Data to fill the appendices is in the Word document on Dropbox. --MKH ****

For many estimates of elasticity of substitution, the resampling technique indicates high uncertainty. 
Where results are not uncertain, 
we find complementarity among factors of production 
for most economies with the exception of China 
where an indication of substitutability between labor and energy is observed.

That so many estimates of elasticity of substitution exhibit large uncertainty, 
often spanning the range from perfect complements ($\sigma, \, \sigma_1 = 0$) 
to perfect substitutes ($\sigma, \, \sigma_1 = \infty$), 
is a troubling characteristic of the estimated CES coefficients 
that is rarely reported in the literature.
For some countries (IR, ZM), 
estimating any elasticity of substitution at all is problematic and fraught with peril. 
Note that this result is unlikely to be caused by poor-quality data alone. 
Presumably the best data in the world (US, UK, JP) exhibits large uncertainty. 
(See US and UK rows in Figures~\ref{fig:CD_Ternary_Scatter_Graph}  
and~\ref{fig:CES_sigma1_delta1_graph}--\ref{fig:CES_ternary_graph}.)

>>>>

The interpretation of elasticities of substitution 
among factors of production ($k$, $l$, and $e$) is challenging,
because a version of each elasticity of substitution appears three or four times 
among the various nests. 
For example, an elasticity of substitution between capital~($k$) and labor~($l$) 
appears in the following nests: 
($\underline{kl}$), 
($\underline{kl}$)$e$, 
($\underline{l}e$)$\underline{k}$, and 
($e\underline{k}$)$\underline{l}$. 
Tables~\ref{tab:sigma_US}--\ref{tab:sigma_ZM} 
provide interpretation of Figures~\ref{fig:CES_sigma1_delta1_graph}
and~\ref{fig:CES_sigma1_delta1_graph}. 
In the tables,
C (c) indicates strong (weak) complementarity, 
S (s) indicates strong (weak) substitutability, and
$\sim\sim\sim$ ($\sim$) indicates large (small) uncertainty.

**** If we stick with these tables, do we want to use $\sim\sim\sim$ to indicate uncertainty? ****

@@

% This is the original sigma table. 
% It has been replaced by macros that make for more-efficient LaTeX code.
% \begin{table} \caption{Elasticities of substitution (US).} \label{tab:sigma_US} 
%   \begin{center}
%     \begin{tabular}{r l c l c l c} 
%       \toprule
%          & \multicolumn{2}{c}{$k$ \& $l$} & \multicolumn{2}{c}{$k$ \& $e$} & \multicolumn{2}{c}{$l$ \& $e$} \\
%       \midrule
%       \multirow{2}{*}{$\sigma_1$} & ($\underline{kl}$) & C 
%                                         &  & 
%                                               & &  \\
%                                   & ($\underline{kl}$)$e$ & C 
%                                         & ($\underline{ek}$)$l$ & $\sim\sim\sim$ 
%                                               & ($\underline{le}$)$k$ & $\sim\sim\sim$ \\
%       \midrule
%       \multirow{3}{*}{$\sigma$} & &
%                                       & ($\underline{k}l$)$\underline{e}$ & $\sim\sim\sim$
%                                             & ($k\underline{l}$)$\underline{e}$ & $\sim\sim\sim$ \\
%                                 & ($\underline{l}e$)$\underline{k}$ & C
%                                       & ($l\underline{e}$)$\underline{k}$ & C
%                                             &  & \\
%                                 & ($e\underline{k}$)$\underline{l}$ & C
%                                       & &
%                                             & ($\underline{e}k$)$\underline{l}$ & C \\
%       \midrule
%                                 & \multicolumn{2}{c}{C} & \multicolumn{2}{c}{C} & \multicolumn{2}{c}{C} \\
%       \bottomrule
%     \end{tabular}
%   \end{center}
% \end{table}

@@

% label is tab:sigma_SA
\eostable{SA}
          {\firstrow{$\sim\sim\sim$}{$\sim\sim\sim$}{C}{C}}
          {\secondrow{C}{C}{$\sim\sim\sim$}{$\sim\sim\sim$}{$\sim\sim\sim$}{$\sim\sim\sim$}}
          {\thirdrow{$\sim\sim\sim$}{C}{C}}

% label is tab:sigma_IR
\eostable{IR}
          {\firstrow{$\sim\sim\sim$}{$\sim\sim\sim$}{$\sim\sim\sim$}{$\sim\sim\sim$}}
          {\secondrow{$\sim\sim\sim$}{$\sim\sim\sim$}{$\sim\sim\sim$}{$\sim\sim\sim$}{C}{C}}
          {\thirdrow{$\sim\sim\sim$}{$\sim\sim\sim$}{$\sim\sim\sim$}}

% label is tab:sigma_TZ
\eostable{TZ}
          {\firstrow{C}{C}{$\sim\sim\sim$}{C}}
          {\secondrow{$\sim\sim\sim$}{$\sim\sim\sim$}{C}{C}{C}{C}}
          {\thirdrow{C}{$\sim\sim\sim$}{C}}

% label is tab:sigma_ZM
\eostable{ZM}
          {\firstrow{$\sim\sim\sim$}{$\sim\sim\sim$}{$\sim\sim\sim$}{$\sim\sim\sim$}}
          {\secondrow{$\sim\sim\sim$}{$\sim\sim\sim$}{$\sim\sim\sim$}
                     {$\sim\sim\sim$}{$\sim\sim\sim$}{$\sim\sim\sim$}}
          {\thirdrow{$\sim\sim\sim$}{$\sim\sim\sim$}{$\sim\sim\sim$}}

For many estimates of elasticity of substitution~($\sigma$ and $\sigma_1$), 
the resampling technique indicates high uncertainty~($\sim\sim\sim$). 
Where results are not uncertain, 
we find complementarity~(C and c) among factors of production 
for most economies with the exception of China 
where an indication of substitutability between labor and energy is observed~(S and s).

That so many estimates of elasticity of substitution exhibit large uncertainty, 
often spanning the range from perfect complements ($\sigma, \, \sigma_1 = 0$) 
to perfect substitutes ($\sigma, \, \sigma_1 = \infty$), 
is a troubling characteristic of the estimated CES coefficients 
that is rarely reported in the literature.
For some economies (Iran and Zambia), 
estimating any elasticity of substitution at all is problematic and fraught with peril. 
Note that this result is unlikely to be caused by poor-quality data alone. 
Presumably the best data in the world 
(United States, United Kingdom, and Japan) exhibits large uncertainty. 
(See United States and United Kingdom rows in Figures~\ref{fig:CD_Ternary_Scatter_Graph}  
and~\ref{fig:CES_sigma1_delta1_graph}--\ref{fig:CES_ternary_graph}.)
